Spatio-angular fluorescence microscopy\\ I. Basic theory

### 2018-12-17 papers/basic-theory/paper/paper.tex
Minor edits from Rudolf, Patrick, Hari. Remove draft tag.

\title{Spatio-angular fluorescence microscopy\\ I. Basic theory (Draft: \today)}

@@

A small but growing group of microscopists is interested in measuring the
prevailing orientation and position of ensembles of fluorophores
\cite{vrabioiu2006, mattheyses2010, mehta2016, mcquilken2017, zhanghao2017}.
These techniques typically use polarizers to make multiple measurements of the
same object with different polarizer orientations, then they use a model of the
dipole excitation and emission processes \cite{fourkas2001} to recover the
orientation of fluorophores using pixel-wise arithmetic. Although these studies
do not adopt the scalar or monopole approximations for the angular part of the
problem, they adopt both approximations when they consider the spatial part of
the problem. Existing works either ignore the spatial reconstruction problem
\cite{vrabioiu2006, mattheyses2010, mehta2016, mcquilken2017} or assume that the
spatial and angular reconstruction problems can be solved sequentially
\cite{zhanghao2017}.

@@

In this paper we focus on modeling an extremely simple optical system---a
single-view fluorescence microscope without polarizers. In future papers of this
series we will extend our models to include polarizers and multi-view
microscopes. Additionally, we have restricted this paper to the forward
problem---the mapping between a known object and the data. In future papers we
will consider the inverse problem, and the singular value decomposition (SVD)
will play a central role.

@@

The spatial bandwidth of a fluorescence microscope is well known to be
$\nu_c = \frac{2\text{NA}}{\lambda}$. In other words, we can increase the
spatial resolution of a fluorescence microscope by increasing the NA of the
instrument or by choosing a fluorophore with a shorter emission wavelength.
Similarly, the angular bandwidth of a fluorescence microscope depends on both
the instrument and the sample.

@@

The angular bandwidth is also sample dependent. Monopoles emit light
isotropically so they have an $\ell$-bandwidth of $\ell_c = 0$, while dipoles
have an $\ell$-bandwidth of $\ell_c = 2$, and higher-order excitation and
detection moments will have even higher bandwidths. Multi-photon excitation and
other non-linear methods can also increase the $\ell$-bandwidth
\cite{brasselet2011}.

>>>>

\title{Spatio-angular fluorescence microscopy\\ I. Basic theory}

@@

A small but growing group of microscopists is interested in measuring the
orientation and position of ensembles of fluorophores \cite{vrabioiu2006,
  mattheyses2010, mehta2016, mcquilken2017, zhanghao2017}. These techniques
typically use polarizers to make multiple measurements of the same object with
different polarizer orientations, then they use a model of the dipole excitation
and emission processes \cite{fourkas2001} to recover the orientation of
fluorophores using pixel-wise arithmetic. Although these studies do not adopt
the scalar or monopole approximations for the angular part of the problem, they
adopt both approximations when they consider the spatial part of the problem.
Existing works either ignore the spatial reconstruction problem
\cite{vrabioiu2006, mattheyses2010, mehta2016, mcquilken2017} or assume that the
spatial and angular reconstruction problems can be solved sequentially
\cite{zhanghao2017}.

@@

In this paper we focus on modeling a single-view fluorescence microscope without
polarizers. In future papers of this series we will extend our models to include
polarizers and multi-view microscopes. Additionally, we have restricted this
paper to the forward problem---the mapping between a known object and the data.
In future papers we will consider the inverse problem, and the singular value
decomposition (SVD) will play a central role.

@@

The spatial bandwidth of a fluorescence microscope is well known to be
$\nu_c = \frac{2\text{NA}}{\lambda}$. In other words, we can increase the
spatial resolution of a fluorescence microscope by increasing the NA of the
instrument or by choosing a fluorophore with a shorter emission wavelength.
Similarly, the angular bandwidth of a fluorescence microscope depends on both
the instrument and the choice of fluorophore.

@@

The angular bandwidth is also fluorophore dependent. Monopoles emit light
isotropically so they have an $\ell$-bandwidth of $\ell_c = 0$, while dipoles
have an $\ell$-bandwidth of $\ell_c = 2$, and higher-order excitation and
detection moments will have even higher bandwidths. Multi-photon excitation and
other non-linear methods can also increase the $\ell$-bandwidth
\cite{brasselet2011}.
